%% This is emulateapj reformatting of the AASTEX sample document
%%
\documentclass[iop]{emulateapj}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{landerson@simonsfoundation.org}

%% You can insert a short comment on the title page using the command below.

\slugcomment{Stars are kinda cool again.}

%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%% The left head contains a list of authors,
%% usually a maximum of three (otherwise use et al.).  The right
%% head is a modified title of up to roughly 44 characters.
%% Running heads will not print in the manuscript style.

\shorttitle{Measuring Mo' Betta Distances}
\shortauthors{Anderson et al}

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

\begin{document}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Mo' Betta Gaia Distance Measurements, \\
    Using Gaia + APASS photometry}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, use \\ to force line breaks.

\author{Lauren Anderson and David Hogg\altaffilmark{1}}
\affil{Center for Computational Astrophysics}

\author{Boris Leistedt}
\affil{NYU}

\author{Adrian Price-Whelan}
\affil{Princeton}

\and

\author{Andrew Casey}
\affil{Somewhere in the UK}

\altaffiltext{1}{NYU}

\begin{abstract}
The Gaia TGAS Catalog contains more than 2 millions parallax measurements.
Safe conversion of a noisy parallax into a posterior belief over distances requires inference with a prior.
Usually this prior represents some kind of belief about the Milky Way.
However, there is multi-band photometry for the TGAS stars from imaging surveys.
This imaging is incredibly informative about stellar distances, as is clear from physical models of stars, and also just from looking at the data themselves.
Here we use color information on TGAS stars from [2MASS or APASS] to build a noise-deconvolved empirical prior distribution of stars in color-magnitude space.
This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
We use the Extreme Deconvolution (XD) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
The algorithm is run not in absolute-magnitude space but in a transformed space where the measurement uncertainty is closer to Gaussian.
The XD optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
The method is capitalizing on statistical shrinkage.
Our posterior parallax estimates are more precise than the Gaia catalog outputs by a factor of [XXX] for the median TGAS star; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
We independently validate our distances by looking at members of Milky Way star clusters; for example, M67 is not visible at all in the TGAS parallax estimates, but appears clearly in our posterior parallax estimates.
All our results, including a posterior parallax and distance sampling for [ZZZ] TGAS stars, are available in companion electronic tables.

\end{abstract}


\keywords{stars}


\section{Introduction}

Given the likelihood measurements of the parallax to local stars from Gaia, and using prior information for the distances to these stars built from the data themselves, we make more precise measurements of the distances to local stars in Gaia than the parallax measurements alone. The prior, which is very similar to a color-magnitude diagram of stars, is informative because stars fall on a locus within this space. By deconvolving the set of observed points with their uncertainties we tighten this locus which therefore informs the individual measurements that they are drawn from a tighter distribution. We build this prior using Extreme Deconvolution. Extreme Deconvolution is a gaussian mixture model of data, specifically data that have heteroscedastic uncertainties, a quality that gaussin mixture models alone cannot take into account. This prior does not include any assumptions about the physical distribution of stars within the galaxy, nor does it include any stellar models.

\section{Data}

We use stars cross matched in the TGAS and APASS set of observations. We require there are no NaNs in any APASS photometric band, as well as every band having positive errors. We check that the match between APASS and TGAS is good using extreme APASS-WISE colors. We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from sampling of posterior Gaia distances from Adrain. There is minimal dust to the matched stars. To build our model here we use the Gaia G band magnitude, parallax and APASS B-V color.

\section{Methods}

We build a prior for the distance measurement of Gaia stars using the Gaia measured parallaxes plus APASS photometry to build a color-magnitude-like diagram that serves as the prior for distance measurements of Gaia stars. Instead of making assumptions about the physical distributions of stars in the galaxy, or using stellar models for the modeled HR diagram of stars, we build a prior from the data themselves.


magSN = parallaxSN = 16 (excluding systematic parallax error in tgas)
train gaussian means, mus, Vs on progressively lower SN dataset

\begin{equation}
\label{eq:bayes}
P(\theta|\textbf{E}, \textbf{I}) = P(\theta ) \frac{P(\textbf{E} |\theta,\textbf{I})}{P(\textbf{E})},
\end{equation}


where $\theta$ is the model parameters that generate the evidence $\textbf{E}$, and P($\theta$) is the prior knowledge of the model, and $P(\textbf{E} |\theta,\textbf{I})$ is the likelihood of the evidence, given the model and some other prior assumptions. Our evidence is the parallaxes and G band magnitudes from Gaia, and the B and V band magnitudes from APASS. We build the prior from this evidence using extreme deconvolution


Crossvalidation

\subsection{Formalism} \label{bozomath}

\section{Results}

\section{Discussion}
Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants

Long discussion of dust

Number of gaussians in XD, and the w parameter



\acknowledgments


\appendix

\section{Appendix material}


\begin{thebibliography}{}
\end{thebibliography}

\clearpage

\end{document}



%%
%% End of file `sample.tex'.
